% Example template for using the unmeethesis style
% This example is for a Master's candidate in Mathematics
% It contains examples of front matter and most sections that the
% typical graduate student would need to include
% By: N. Doren 02/10/00
%     Minor mods by N. Doren 08/26/11

% Use the following specification for BOTTOM page numbering:
\documentclass[botnum, fleqn]{unmeethesis}
% OR
% Use the following specification for TOP page numbering:
% \documentclass[fleqn]{unmeethesis}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{minted}

\makeatletter
\def\BState{\State\hskip-\ALG@thistlm}
\makeatother

\begin{document}

  \frontmatter

  % Uncomment the next command if you see weird paragraph spacing:
  % That is, if you see paragraphs float with lots of white space
  % in between them:

  % \setlength{\parskip}{0.30cm}

  \title{Distributed Video Analysis for the  Advancing  \\
  Out of School Learning in Mathematics and Engineering Project}

  \author{Cody Wilson Eilar}
  \degreesubject{M.S., Computer Engineering}
  \degree{Master of Science \\ Computer Engineering}
  \documenttype{Thesis}
  \previousdegrees{B.S., University of New Mexico, 2010}
  \date{July, \thisyear}

  \maketitle

  %\makecopyright
  %Copyright page is no longer necessary D. Murrell


  \begin{acknowledgments}
    \vspace{1.1in}
    I would like to thank my advisor, Professor Marios Pattichis, for his
    continuous support in reviewing my source code and for helping me shape my
    ideas for this thesis. I would also like to thank Elmyra Grelle for helping
    me stay on track with all the necessary requirements to graduate. Finally, I
    would like to thank Sandia National Laboratories for supporting me
    financially in my academic endeavors.
  \end{acknowledgments}

  \maketitleabstract %(required even though there's no abstract title anymore)

  \begin{abstract}
    This thesis proposes a distributed processing system for analyzing
    videos acquired through the advancing out of school learning in
    mathematics and engineering (AOLME) project. The
    proposed architecture is demonstrated by detecting writing and typing
    activities based on the Video Analysis (ViDA) system. VIDA
    leverages Amazon Web Services (AWS) to optimally distribute segments of
    video to a heterogenous compute cloud consisting of machines that have both
    CPU and GPU processing hardware onboard. The master node is responsible for
    distributing the videos and the slave nodes perform feature reduction on the
    videos by calculating a handful of cumulative distribution functions (CDF)
    to then be returned to the master for classification using a support vector
    machine (SVM). This thesis will demonstrate the accuracy,
    scalability and flexibility of ViDA for videos collected through AOLME. Furthermore,
    this thesis delves into the architectural details essential to distributed
    video processing on cloud infrastructures.

    \clearpage %(required for 1-page abstract)
  \end{abstract}

  \tableofcontents
  \listoffigures
  \listoftables

  \begin{glossary}{Longest  string}
    \item[$a_{lm}$]
    Taylor series coefficients, where $l,m = \{0..2\}$
  \end{glossary}

  \mainmatter

  % Include the topic chapters
  \include{introduction}
  \include{activity_classification}
  \include{architecture}

  \chapter*{Appendices}

  \addcontentsline{toc}{chapter}{Appendices}
  % Next lines duplicated from .toc file and used to create mini
  % "Appendix Table of Contents," if desired:
  %\contentsline {chapter}{\numberline {A}Proving $E=MC^2$}{4}
 % \contentsline {chapter}{\numberline {B}Derivation of $A = \pi r^2$}{5}
  % End mini table of contents

  \appendix
  \include{cents_and_orients}

  \bibliographystyle{acm}
  \bibliography{ivpcl_publications}

\end{document}
